---
title: "Every number carries the method that produced it"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Every number carries the method that produced it}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Ten published ways of computing one jump height disagree by 3.51 cm on 244 real trials.
The training effect the source study was built to detect is 1.98 cm. plateforce records
which one you used.

```{r setup}
library(plateforce)
```

## A trace, and the rate it was recorded at

A trial is a vector of vertical ground reaction force in newtons, and the rate it was
recorded at. The rate has no default: guessing it scales every velocity, displacement,
impulse and rate of force development by the ratio of the two rates.

```{r trial}
trial <- pf_trial(rep(700, 1200), sample_rate_hz = 1200)
trial
```

Reading from a file works the same way, through `pf_read_force_file()`, which states back
the delimiter it used, the column it read as force, the rows it read and the samples the
sentinel convention treated as missing.

## Naming the three rules

An analysis names the rule that establishes system weight, the rule that places the start
of the jump, and the rule that places the instant of takeoff. Each is a registry
identifier.

```{r bindings}
head(pf_bindings(), 4)
```

```{r analyse}
result <- analyse_countermovement_jump(
  trial,
  weighing = "bwepoch.fixed_window",
  onset = "onset.threshold.noise_relative",
  takeoff = "takeoff.threshold.absolute_force"
)
```

This trace is a person standing still, so the landmark rules have nothing to find, and
each one says so by name and by parameter rather than returning a number anyway.

```{r warnings}
result@warnings
```

## Reading a value, and what produced it

```{r value}
weight <- pf_value(result, "system_weight_newtons")
weight
```

The number is `@value`. Everything that produced it is `@provenance`, and the rules
upstream of it are `@depends_on`, each with the parameters it read and where each value
came from.

```{r provenance}
chain <- weight@provenance@depends_on
chain[[1]]@method_id
chain[[1]]@parameters
```

`assumed` means the rule supplied that value because the caller did not. State one and it
is recorded as stated:

```{r stated}
stated <- analyse_countermovement_jump(
  trial,
  weighing = "bwepoch.fixed_window",
  weighing_parameters = list(duration = 0.5),
  onset = "onset.threshold.noise_relative",
  takeoff = "takeoff.threshold.absolute_force"
)
bound <- pf_value(stated, "system_weight_newtons")@provenance@depends_on[[1]]@parameters
bound[bound[["name"]] == "duration", ]
```

Reading a shortened property name is an error rather than a neighbouring property's value,
and putting the object into arithmetic asks you to name the number:

```{r refuse-arithmetic, error = TRUE}
weight + 1
```

```{r value-out}
weight@value
```

## When a rule declines

A rule that cannot answer raises a condition classed by its code, so you branch on the
fields rather than on the sentence.

```{r refusal}
condition <- tryCatch(
  analyse_countermovement_jump(
    trial,
    weighing = "bwepoch.fixed_window",
    onset = "onset.not.a.rule",
    takeoff = "takeoff.threshold.absolute_force"
  ),
  plateforce_refusal = identity
)
class(condition)
condition[["code"]]
```

Read a field with `[[`, which matches the whole name. `condition$method` is an error here,
because `$` on a list matches a shortened name without saying so.

## The registry

```{r registry}
registry <- pf_registry()
registry@census
```

The two populations are counted apart and never added together, and each derived count
sits beside the population it was taken over. One entry reads out in full:

```{r entry}
entry <- pf_entry("onset.threshold.noise_relative")
entry[["construct"]]
entry[["title"]]
```
